---
title: "WGEA Dataset Analysis - Combined Dataset"
author: "Analysis Team"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(readr)
library(janitor)
```

## Loading and Combining WGEA Datasets

This document combines all WGEA datasets based on the Australian Business Number (ABN) and removes duplicate features.

```{r load-data}
# Load all datasets
action_on_gender <- read_csv("wgea_public_dataset_2024/wgea_questionnaire_action_on_gender_equality_2024.csv")
employee_support <- read_csv("wgea_public_dataset_2024/wgea_questionnaire_employee_support_2024.csv")
flexible_work <- read_csv("wgea_public_dataset_2024/wgea_questionnaire_flexible_work_2024.csv")
harm_prevention <- read_csv("wgea_public_dataset_2024/wgea_questionnaire_harm_prevention_2024.csv")
workplace_overview <- read_csv("wgea_public_dataset_2024/wgea_questionnaire_workplace_overview_2024.csv")
workforce_composition <- read_csv("wgea_public_dataset_2024/wgea_workforce_composition_2024.csv")
workforce_management <- read_csv("wgea_public_dataset_2024/wgea_workforce_management_statistics_2024.csv")

# Note: wgea_questionnaire_catalogue_2024.csv is a reference table without ABN, so we'll handle it separately
catalogue <- read_csv("wgea_public_dataset_2024/wgea_questionnaire_catalogue_2024.csv")

print("Datasets loaded successfully!")
```

```{r examine-structure}
# Examine the structure of each dataset
cat("Action on Gender Equality dimensions:", dim(action_on_gender), "\n")
cat("Employee Support dimensions:", dim(employee_support), "\n")
cat("Flexible Work dimensions:", dim(flexible_work), "\n")
cat("Harm Prevention dimensions:", dim(harm_prevention), "\n")
cat("Workplace Overview dimensions:", dim(workplace_overview), "\n")
cat("Workforce Composition dimensions:", dim(workforce_composition), "\n")
cat("Workforce Management dimensions:", dim(workforce_management), "\n")
cat("Catalogue dimensions:", dim(catalogue), "\n")
```

```{r identify-common-columns}
# Identify common columns across datasets (excluding catalogue)
datasets_with_abn <- list(
  action_on_gender = action_on_gender,
  employee_support = employee_support,
  flexible_work = flexible_work,
  harm_prevention = harm_prevention,
  workplace_overview = workplace_overview,
  workforce_composition = workforce_composition,
  workforce_management = workforce_management
)

# Get column names for each dataset
column_names <- lapply(datasets_with_abn, names)

# Find common columns across all datasets
common_cols <- Reduce(intersect, column_names)
cat("Common columns across all datasets:\n")
print(common_cols)

# Find unique columns in each dataset
unique_cols <- lapply(names(datasets_with_abn), function(dataset_name) {
  cols <- column_names[[dataset_name]]
  unique_cols <- setdiff(cols, common_cols)
  cat("\nUnique columns in", dataset_name, ":\n")
  print(unique_cols)
  return(unique_cols)
})
names(unique_cols) <- names(datasets_with_abn)
```

```{r create-base-dataset}
# Create a base dataset with unique employer information
# Using workforce_composition as it has the most comprehensive employer info
base_employer_info <- workforce_composition %>%
  select(all_of(common_cols)) %>%
  distinct() %>%
  arrange(employer_abn)

cat("Base employer information created with", nrow(base_employer_info), "unique employers\n")
```

```{r combine-questionnaire-data}
# Combine all questionnaire datasets
# These have similar structure with questions and responses

questionnaire_datasets <- list(
  action_on_gender = action_on_gender,
  employee_support = employee_support,
  flexible_work = flexible_work,
  harm_prevention = harm_prevention,
  workplace_overview = workplace_overview
)

# Add a source column to track which dataset each record came from
questionnaire_combined <- bind_rows(
  action_on_gender %>% mutate(data_source = "action_on_gender"),
  employee_support %>% mutate(data_source = "employee_support"),
  flexible_work %>% mutate(data_source = "flexible_work"),
  harm_prevention %>% mutate(data_source = "harm_prevention"),
  workplace_overview %>% mutate(data_source = "workplace_overview")
)

cat("Combined questionnaire data has", nrow(questionnaire_combined), "rows\n")
```

```{r process-workforce-data}
# Process workforce composition data
workforce_comp_processed <- workforce_composition %>%
  mutate(data_source = "workforce_composition")

# Process workforce management data  
workforce_mgmt_processed <- workforce_management %>%
  mutate(data_source = "workforce_management")

cat("Workforce composition data:", nrow(workforce_comp_processed), "rows\n")
cat("Workforce management data:", nrow(workforce_mgmt_processed), "rows\n")
```

```{r create-combined-dataset}
# Create the final combined dataset
# Since the data structures are quite different, we'll create separate summary views

# 1. Employer-level summary combining all unique employer attributes
employer_summary <- base_employer_info %>%
  group_by(employer_abn, corporate_group_name, employer_name, employer_size, 
           anzsic_code, anzsic_division, anzsic_subdivision, anzsic_group, anzsic_class) %>%
  summarise(
    reporting_year = first(reporting_year),
    is_relevant_employer = first(is_relevant_employer),
    .groups = "drop"
  )

# 2. Add questionnaire response counts by employer
questionnaire_summary <- questionnaire_combined %>%
  group_by(employer_abn) %>%
  summarise(
    total_questionnaire_responses = n(),
    action_on_gender_responses = sum(data_source == "action_on_gender"),
    employee_support_responses = sum(data_source == "employee_support"),
    flexible_work_responses = sum(data_source == "flexible_work"),
    harm_prevention_responses = sum(data_source == "harm_prevention"),
    workplace_overview_responses = sum(data_source == "workplace_overview"),
    .groups = "drop"
  )

# 3. Add workforce composition summary
workforce_comp_summary <- workforce_comp_processed %>%
  group_by(employer_abn) %>%
  summarise(
    total_employees = sum(n_employees, na.rm = TRUE),
    unique_occupations = n_distinct(occupation, na.rm = TRUE),
    unique_manager_categories = n_distinct(manager_category, na.rm = TRUE),
    male_employees = sum(n_employees[gender == "Male"], na.rm = TRUE),
    female_employees = sum(n_employees[gender == "Female"], na.rm = TRUE),
    .groups = "drop"
  )

# 4. Add workforce management summary
workforce_mgmt_summary <- workforce_mgmt_processed %>%
  group_by(employer_abn) %>%
  summarise(
    total_workforce_movements = sum(n_employees, na.rm = TRUE),
    unique_movement_types = n_distinct(movement_type, na.rm = TRUE),
    unique_manager_types = n_distinct(manager_type, na.rm = TRUE),
    .groups = "drop"
  )

# Combine all summaries
combined_dataset <- employer_summary %>%
  left_join(questionnaire_summary, by = "employer_abn") %>%
  left_join(workforce_comp_summary, by = "employer_abn") %>%
  left_join(workforce_mgmt_summary, by = "employer_abn") %>%
  # Replace NA values with 0 for count variables
  mutate(across(ends_with("_responses") | starts_with("total_") | starts_with("unique_") | 
                ends_with("_employees"), ~ ifelse(is.na(.), 0, .)))

cat("Final combined dataset created with", nrow(combined_dataset), "employers and", ncol(combined_dataset), "variables\n")
```

```{r examine-final-dataset}
# Examine the final combined dataset
cat("Final dataset structure:\n")
str(combined_dataset)

cat("\nFirst few rows of the combined dataset:\n")
head(combined_dataset)

cat("\nSummary statistics:\n")
summary(combined_dataset %>% select_if(is.numeric))
```

```{r save-combined-dataset}
# Save the combined dataset
write_csv(combined_dataset, "wgea_combined_dataset_2024.csv")

cat("Combined dataset saved as 'wgea_combined_dataset_2024.csv'\n")
cat("Dataset contains", nrow(combined_dataset), "unique employers with comprehensive information\n")
```

```{r data-quality-checks}
# Perform data quality checks
cat("Data Quality Summary:\n")
cat("- Total unique employers:", n_distinct(combined_dataset$employer_abn), "\n")
cat("- Employers with questionnaire responses:", sum(combined_dataset$total_questionnaire_responses > 0), "\n")
cat("- Employers with workforce composition data:", sum(combined_dataset$total_employees > 0), "\n")
cat("- Employers with workforce management data:", sum(combined_dataset$total_workforce_movements > 0), "\n")

# Check for any duplicate ABNs (should be 0)
duplicates <- combined_dataset %>%
  group_by(employer_abn) %>%
  summarise(count = n()) %>%
  filter(count > 1)

cat("- Duplicate ABNs in final dataset:", nrow(duplicates), "\n")

# Display column names and data types
cat("\nFinal dataset columns:\n")
combined_dataset %>% 
  summarise_all(class) %>% 
  gather(variable, class) %>%
  print()
```

```{r feature-engineering-for-classification}
# Feature Engineering for Female Management Representation Classification

# First, let's create detailed management and gender statistics from workforce composition
management_gender_stats <- workforce_composition %>%
  filter(!is.na(manager_category) & manager_category != "" & 
         !is.na(gender) & gender %in% c("Male", "Female")) %>%
  group_by(employer_abn, manager_category, gender) %>%
  summarise(n_managers = sum(n_employees, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = gender, values_from = n_managers, values_fill = 0) %>%
  mutate(
    total_managers_by_category = Male + Female,
    female_proportion_by_category = ifelse(total_managers_by_category > 0, 
                                         Female / total_managers_by_category, 0)
  )

# Aggregate management statistics by employer
employer_management_stats <- management_gender_stats %>%
  group_by(employer_abn) %>%
  summarise(
    total_male_managers = sum(Male),
    total_female_managers = sum(Female),
    total_managers = sum(total_managers_by_category),
    avg_female_mgmt_proportion = mean(female_proportion_by_category, na.rm = TRUE),
    weighted_female_mgmt_proportion = ifelse(total_managers > 0, 
                                           total_female_managers / total_managers, 0),
    mgmt_categories_count = n_distinct(manager_category),
    # Management diversity metrics
    mgmt_categories_with_females = sum(Female > 0),
    mgmt_categories_female_majority = sum(female_proportion_by_category > 0.5),
    mgmt_gender_balance_score = 1 - abs(0.5 - weighted_female_mgmt_proportion) * 2,
    .groups = "drop"
  )

# Create the target variable: Female management representation categories
employer_management_stats <- employer_management_stats %>%
  mutate(
    female_mgmt_category = case_when(
      total_managers == 0 ~ "No_Management_Data",
      weighted_female_mgmt_proportion < 0.4 ~ "Low_Female_Representation",
      weighted_female_mgmt_proportion >= 0.4 & weighted_female_mgmt_proportion <= 0.6 ~ "Balanced_Representation", 
      weighted_female_mgmt_proportion > 0.6 ~ "High_Female_Representation",
      TRUE ~ "Unknown"
    )
  )

cat("Target variable distribution:\n")
table(employer_management_stats$female_mgmt_category)
```

```{r advanced-feature-engineering}
# Create additional features for the classification model

# 1. Industry and size features
industry_size_features <- combined_dataset %>%
  select(employer_abn, employer_size, anzsic_division, anzsic_subdivision) %>%
  mutate(
    # Encode employer size as ordinal
    size_numeric = case_when(
      employer_size == "Small (100-199)" ~ 1,
      employer_size == "Medium (200-999)" ~ 2,  
      employer_size == "Large (1000+)" ~ 3,
      TRUE ~ 0
    ),
    # Create industry groups
    industry_group = case_when(
      anzsic_division %in% c("Education and Training", "Health Care and Social Assistance") ~ "Education_Health",
      anzsic_division %in% c("Public Administration and Safety", "Administrative and Support Services") ~ "Government_Admin",
      anzsic_division %in% c("Financial and Insurance Services", "Professional, Scientific and Technical Services") ~ "Finance_Professional",
      anzsic_division %in% c("Manufacturing", "Construction", "Mining") ~ "Industrial",
      anzsic_division %in% c("Retail Trade", "Accommodation and Food Services", "Arts and Recreation Services") ~ "Service_Consumer",
      TRUE ~ "Other"
    )
  )

# 2. Workforce diversity features
workforce_diversity_features <- workforce_composition %>%
  group_by(employer_abn) %>%
  summarise(
    # Overall workforce gender balance
    total_workforce = sum(n_employees, na.rm = TRUE),
    female_workforce_prop = sum(n_employees[gender == "Female"], na.rm = TRUE) / 
                           sum(n_employees, na.rm = TRUE),
    male_workforce_prop = sum(n_employees[gender == "Male"], na.rm = TRUE) / 
                         sum(n_employees, na.rm = TRUE),
    
    # Occupational diversity
    unique_occupations = n_distinct(occupation, na.rm = TRUE),
    occupation_diversity_index = 1 - sum((n_employees/sum(n_employees))^2, na.rm = TRUE),
    
    # Employment type diversity
    full_time_prop = sum(n_employees[employment_type == "Full-time"], na.rm = TRUE) / 
                    sum(n_employees, na.rm = TRUE),
    part_time_prop = sum(n_employees[employment_type == "Part-time"], na.rm = TRUE) / 
                    sum(n_employees, na.rm = TRUE),
    
    # Gender distribution across employment types
    female_fulltime_prop = sum(n_employees[gender == "Female" & employment_type == "Full-time"], na.rm = TRUE) / 
                          sum(n_employees[gender == "Female"], na.rm = TRUE),
    female_parttime_prop = sum(n_employees[gender == "Female" & employment_type == "Part-time"], na.rm = TRUE) / 
                          sum(n_employees[gender == "Female"], na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  mutate(
    # Replace NaN and Inf with 0
    across(everything(), ~ifelse(is.nan(.) | is.infinite(.), 0, .)),
    
    # Workforce size categories
    workforce_size_category = case_when(
      total_workforce < 200 ~ "Small_Workforce",
      total_workforce >= 200 & total_workforce < 1000 ~ "Medium_Workforce",
      total_workforce >= 1000 ~ "Large_Workforce",
      TRUE ~ "Unknown_Size"
    ),
    
    # Gender balance score (closer to 1 = more balanced)
    workforce_gender_balance = 1 - abs(0.5 - female_workforce_prop) * 2
  )

# 3. Policy and initiative features from questionnaire data
policy_features <- questionnaire_combined %>%
  group_by(employer_abn, data_source, question_text) %>%
  summarise(
    response_value = first(response),
    .groups = "drop"
  ) %>%
  # Focus on key gender equality indicators
  filter(
    str_detect(tolower(question_text), "gender|equality|diversity|flexible|parental|leadership|promotion|mentoring|target") |
    str_detect(tolower(response), "yes|no|implemented|available")
  ) %>%
  group_by(employer_abn) %>%
  summarise(
    # Count of positive policy responses
    total_policy_questions = n(),
    positive_responses = sum(str_detect(tolower(response_value), "yes|implemented|available|established"), na.rm = TRUE),
    policy_engagement_score = ifelse(total_policy_questions > 0, 
                                   positive_responses / total_policy_questions, 0),
    
    # Specific policy areas
    has_gender_targets = any(str_detect(tolower(question_text), "target") & 
                           str_detect(tolower(response_value), "yes|implemented")),
    has_flexible_work = any(str_detect(tolower(question_text), "flexible") & 
                          str_detect(tolower(response_value), "yes|available")),
    has_mentoring = any(str_detect(tolower(question_text), "mentor") & 
                       str_detect(tolower(response_value), "yes|implemented")),
    has_leadership_programs = any(str_detect(tolower(question_text), "leadership|development") & 
                                str_detect(tolower(response_value), "yes|available")),
    
    .groups = "drop"
  ) %>%
  mutate(
    # Convert logical to numeric
    across(starts_with("has_"), ~as.numeric(.))
  )

# 4. Workforce movement features
movement_features <- workforce_management %>%
  group_by(employer_abn, movement_type, gender) %>%
  summarise(movement_count = sum(n_employees, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = c(movement_type, gender), 
              values_from = movement_count, values_fill = 0,
              names_sep = "_") %>%
  mutate(
    # Calculate promotion rates by gender
    total_promotions = rowSums(select(., contains("Promotion")), na.rm = TRUE),
    female_promotions = rowSums(select(., contains("Promotion_Female")), na.rm = TRUE),
    male_promotions = rowSums(select(., contains("Promotion_Male")), na.rm = TRUE),
    
    female_promotion_ratio = ifelse(total_promotions > 0, 
                                  female_promotions / total_promotions, 0),
    
    # Calculate turnover patterns
    total_resignations = rowSums(select(., contains("Resignation")), na.rm = TRUE),
    female_resignations = rowSums(select(., contains("Resignation_Female")), na.rm = TRUE),
    
    female_turnover_ratio = ifelse(total_resignations > 0,
                                 female_resignations / total_resignations, 0)
  ) %>%
  select(employer_abn, total_promotions, female_promotion_ratio, 
         total_resignations, female_turnover_ratio)

cat("Feature engineering completed successfully!\n")
```

```{r create-final-modeling-dataset}
# Combine all features into final modeling dataset
modeling_dataset <- employer_management_stats %>%
  left_join(industry_size_features, by = "employer_abn") %>%
  left_join(workforce_diversity_features, by = "employer_abn") %>%
  left_join(policy_features, by = "employer_abn") %>%
  left_join(movement_features, by = "employer_abn") %>%
  # Add original combined dataset features
  left_join(combined_dataset %>% 
           select(employer_abn, total_questionnaire_responses, 
                  action_on_gender_responses, employee_support_responses,
                  flexible_work_responses, harm_prevention_responses,
                  workplace_overview_responses), 
           by = "employer_abn") %>%
  # Clean up missing values
  mutate(
    across(where(is.numeric), ~ifelse(is.na(.), 0, .)),
    across(where(is.logical), ~ifelse(is.na(.), FALSE, .)),
    industry_group = ifelse(is.na(industry_group), "Other", industry_group),
    workforce_size_category = ifelse(is.na(workforce_size_category), "Unknown_Size", workforce_size_category)
  ) %>%
  # Remove companies with no management data for modeling
  filter(female_mgmt_category != "No_Management_Data")

# Create additional derived features
modeling_dataset <- modeling_dataset %>%
  mutate(
    # Interaction features
    size_industry_interaction = paste(workforce_size_category, industry_group, sep = "_"),
    policy_size_interaction = policy_engagement_score * size_numeric,
    diversity_balance_score = (workforce_gender_balance + mgmt_gender_balance_score) / 2,
    
    # Categorical versions of continuous variables for potential use
    female_workforce_category = cut(female_workforce_prop, 
                                  breaks = c(0, 0.3, 0.7, 1), 
                                  labels = c("Low_Female", "Balanced", "High_Female"),
                                  include.lowest = TRUE),
    
    policy_engagement_category = cut(policy_engagement_score,
                                   breaks = c(0, 0.3, 0.7, 1),
                                   labels = c("Low_Engagement", "Medium_Engagement", "High_Engagement"),
                                   include.lowest = TRUE)
  )

# Display the final dataset structure
cat("Final modeling dataset created!\n")
cat("Dimensions:", dim(modeling_dataset), "\n")
cat("Target variable distribution:\n")
print(table(modeling_dataset$female_mgmt_category))

cat("\nKey features summary:\n")
feature_columns <- setdiff(names(modeling_dataset), 
                          c("employer_abn", "female_mgmt_category"))
cat("Total features available:", length(feature_columns), "\n")

# Show feature importance preview
numeric_features <- modeling_dataset %>% 
  select(where(is.numeric)) %>%
  select(-employer_abn) %>%
  names()

cat("Numeric features (", length(numeric_features), "):\n")
print(numeric_features)

categorical_features <- modeling_dataset %>%
  select(where(function(x) is.character(x) | is.factor(x))) %>%
  select(-employer_abn, -female_mgmt_category) %>%
  names()

cat("\nCategorical features (", length(categorical_features), "):\n") 
print(categorical_features)
```

```{r save-modeling-dataset}
# Save the modeling dataset
write_csv(modeling_dataset, "wgea_modeling_dataset_2024.csv")

# Create a feature description file
feature_descriptions <- data.frame(
  Feature_Name = names(modeling_dataset),
  Feature_Type = sapply(modeling_dataset, function(x) class(x)[1]),
  Description = c(
    "Australian Business Number (identifier)",
    "Number of male managers", 
    "Number of female managers",
    "Total number of managers",
    "Average female proportion across management categories",
    "Weighted female proportion in management (target calculation base)",
    "Number of management categories",
    "Management categories with at least one female",
    "Management categories with female majority",
    "Management gender balance score (0-1, 1 = perfectly balanced)",
    "Target variable: Female management representation level",
    "Employer size as numeric (1=Small, 2=Medium, 3=Large)",
    "ANZSIC industry division",
    "ANZSIC industry subdivision", 
    "Grouped industry categories",
    "Total workforce size",
    "Proportion of female employees in workforce",
    "Proportion of male employees in workforce",
    "Number of unique occupations",
    "Occupation diversity index",
    "Proportion of full-time employees",
    "Proportion of part-time employees", 
    "Proportion of female employees who are full-time",
    "Proportion of female employees who are part-time",
    "Workforce size category",
    "Overall workforce gender balance score",
    "Total policy-related questionnaire responses",
    "Number of positive policy responses",
    "Policy engagement score (positive responses / total)",
    "Has gender equality targets",
    "Has flexible work arrangements",
    "Has mentoring programs", 
    "Has leadership development programs",
    "Total promotions across all employees",
    "Female proportion of promotions",
    "Total resignations across all employees",
    "Female proportion of resignations",
    paste("Other features from combined dataset:", 
          paste(setdiff(names(modeling_dataset), names(feature_descriptions)[1:37]), collapse = ", "))
  )[1:length(names(modeling_dataset))]
)

write_csv(feature_descriptions, "feature_descriptions.csv")

cat("Modeling dataset saved as 'wgea_modeling_dataset_2024.csv'\n")
cat("Feature descriptions saved as 'feature_descriptions.csv'\n")
cat("Ready for classification modeling!\n")
```

## Summary

This analysis has successfully combined all WGEA datasets and engineered comprehensive features for predicting female management representation. The approach included:

### Target Variable Creation:
- **Low_Female_Representation**: < 40% female managers
- **Balanced_Representation**: 40-60% female managers  
- **High_Female_Representation**: > 60% female managers

### Feature Categories Engineered:

1. **Management Features**: Total managers, gender proportions, balance scores
2. **Industry & Size Features**: Company size, industry groupings, size-industry interactions
3. **Workforce Diversity Features**: Gender balance, occupational diversity, employment patterns
4. **Policy & Initiative Features**: Gender equality policies, flexible work, mentoring programs
5. **Workforce Movement Features**: Promotion patterns, turnover rates by gender
6. **Engagement Features**: Questionnaire response patterns, policy engagement scores

### Key Features for Classification:
- `weighted_female_mgmt_proportion` (target calculation base)
- `workforce_gender_balance` (overall gender balance)
- `policy_engagement_score` (commitment to gender initiatives)
- `industry_group` (industry context)
- `size_numeric` (company size effect)
- `female_promotion_ratio` (career advancement patterns)
- Various interaction terms and derived features

The final modeling dataset contains **comprehensive features** engineered specifically for predicting female management representation, ready for machine learning algorithms!